\propertitle

### 2019-11-18 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T10:30:55+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
log-ratio of the likelihoods of the hypotheses $M''$ and $M'$ in view of
the data $\yf$:
\begin{equation}
  \label{eq:bayes_factor}
  \ydi(M'',M') =
  \log\bigl[\pf(\yf \| M'')/
  \pf(\yf \| M')\bigr].
\end{equation}
This is called their \emph{relative weight of evidence}, the logarithm of
their \emph{relative Bayes factor}
\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
this in appendix**. So the exponential (same base) of $\ydi(M'',M')$ tells
us how much more probable the data $\yf$ are, in view of $M''$ vs $M'$. If
we also introduce the pre-data probabilities of the two hypotheses we
obtain their probabilities conditional on the data.

@@

Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
in the first two, four, five moments. For a full network of size
$N=10\,000$, we obtain the following differences:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
\\    \ydi(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}. In words, the measured frequencies of the
sample activity are 35 orders of magnitude more probable under sufficiency
of the first four moments are sufficient, than under sufficiency of the
first two only; but they are about equally probable ($10^{0.016} = 1.04$)
whether we consider the first five moments to be sufficient or just the
first four.

Compare the results above with those obtained by applying the method at the
sample level only:
\begin{equation}
  \label{eq:diff_suff_n}
  \begin{split}
  \ydi_{\text{sample level}}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
\\    \ydi_{\text{sample level}}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart,
\end{split}
\end{equation}
that is, the data are 1\,220 orders of magnitude more probable conditional
on four moments as opposed to two, and $10^{1.4}=25$ times more probable
conditional on five moments as opposed to five.

The application at the full-network level therefore leads to different
conclusions about the relative importance of the moments from the
application at the sample level. It is also possible to use the
measure~\eqref{eq:measure_suffic} and its probabilistic
meaninig~\eqref{eq:bayes_factor} to compare the probability of the data
$\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
the full-network level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
5-moment sufficiency at the sample level, $N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi(M_{5, N}, M_{5, n}) = 35\;\nat = 15\;\hart,
\end{equation}
that is, the data are 15 orders of magnitude more likely under the first
hypothesis than under the second.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T10:37:30+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Why is this a natural measure? Because $\ydi(M'',M')$ is equal to the
log-ratio of the likelihoods of the hypotheses $M''$ and $M'$ in view of
the data $\yf$:
\begin{equation}
  \label{eq:bayes_factor}
  \ydi(M'',M') =
  \log\bigl[\pf(\yf \| M'')/
  \pf(\yf \| M')\bigr].
\end{equation}
This is called their \emph{relative weight of evidence}, the logarithm of
their \emph{relative Bayes factor}
\citep[\chap~6]{good1950}{good1975,good1981,good1985,good1983}[\sect~1.4]{osteyeeetal1974}{mackay1992,kassetal1995}[see
also][p.~421]{jeffreys1936}[\chaps~V, VI, A]{jeffreys1939_r1983}. We prove
this equality in appendix**. So the exponential of $\ydi(M'',M')$ tells
us how much more probable the data $\yf$ are, conditional on $M''$ versus $M'$. If
we assign the pre-data probabilities of the two hypotheses, this measure
also gives us the ratio of their probabilities conditional on the data.

@@

Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
in the first two, four, five moments. For a full network of size
$N=10\,000$, we obtain the following differences:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
\\    \ydi(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}. In words, the measured frequencies of the
sample activity are 35 orders of magnitude more probable under sufficiency
of the first four moments than under sufficiency of the first two only; but
they are about equally probable ($10^{0.016} = 1.04$) whether we consider
the first five moments to be sufficient or just the first four.

Compare the results above with those obtained by applying the method at the
sample level only:
\begin{equation}
  \label{eq:diff_suff_n}
  \begin{split}
  \ydi_{\text{sample level}}(M_{4},M_{2}) &= 280\;\nat = 1\,220\;\hart,
\\    \ydi_{\text{sample level}}(M_{5},M_{4}) &= 3.3\;\nat = 1.4\;\hart,
\end{split}
\end{equation}
that is, the data are 1\,220 orders of magnitude more probable conditional
on four moments than conditional on two, and $10^{1.4}=25$ times more
probable conditional on five moments than on four moments.

The application at the full-network level therefore leads to different
conclusions about the relative importance of the moments from the
application at the sample level. It is also possible to use the
measure~\eqref{eq:measure_suffic} and its probabilistic
meaning~\eqref{eq:bayes_factor} to compare the probability of the data
$\yf$ conditional on the hypothesis $M_{5, N}$ of 5-moment sufficiency at
the full-network level, $N=10\,000$, versus the hypothesis $M_{5, n}$ of
5-moment sufficiency at the sample level, $N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi(M_{5, N}, M_{5, n}) = 35\;\nat = 15\;\hart,
\end{equation}
that is, the data are 15 orders of magnitude more likely under the first
hypothesis than under the second.
